---
title: "Lab5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question 1 - produce F value that is lower than the critical value

```{r}

library(tibble)
library(ggplot2)

levels <- 3
n_per_level <- 10

critical_value<-qf(.95,2,27)

for(i in 1:1000){
random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rnorm(levels*n_per_level, 0, 1))
aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]


if (simulated_F<critical_value) break}

summary(aov.out)

ggplot(random_data,aes(x=IV,y=DV))+geom_bar(stat="summary",fun="mean")+geom_point()

```

Since we are simulating a dataset that could be produced by the null hypothesis, with an F value that is smaller than the critical value, we therefore accept the null, and we would be correct in doing so. 

##Question 2 - produce F value that is higher than the critical value

```{r}

library(tibble)
library(ggplot2)

levels <- 3
n_per_level <- 10

critical_value<-qf(.95,2,27)

for(i in 1:1000){
random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rnorm(levels*n_per_level, 0, 1))
aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]


if (simulated_F>critical_value) break}

simulated_F>critical_value

summary(aov.out)

ggplot(random_data,aes(x=IV,y=DV))+geom_bar(stat="summary",fun="mean")+geom_point()

```

Since we are simulating a dataset that could be produced by the null hypothesis, with an F value that is higher than the critical value, we therefore reject the null, however we would be incorrect in doing so - this is a type I error. 

##Question 3

```{r}
library(tibble)
library(ggplot2)

levels <- 3
n_per_level <- 10

save_F_values <- length(1000)

for(i in 1:1000){
random_data <- tibble(subjects = 1:(levels*n_per_level),
                      IV = as.factor(rep(1:levels, each = n_per_level)),
                      DV = rt(levels*n_per_level, 1)
                      )
aov.out <- aov(DV ~ IV, data = random_data)
simulated_F <- summary(aov.out)[[1]]$`F value`[1]
save_F_values[i] <- simulated_F
}



F_comparison <- tibble(type = rep(c("analytic","simulated_rt"), each = 1000),
                        F_value = c(rf(1000,levels-1,(levels*n_per_level)-levels),save_F_values))

ggplot(F_comparison, aes(x=F_value, color=type))+
  geom_freqpoly(bins=50)


```
